<!--- Title Slide --->

```{r}
#| label: setup
#| include: false
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width = 7,
  fig.height = 5,
  fig.align = "center",
  out.width = "100%"
)

# Load required packages for QR code generation
if (!require("qrcode")) install.packages("qrcode")
library(qrcode)

if (!require("here")) install.packages("here")
library(here)
here::i_am("presentation.qmd")

include_local_figure <- function(data_source) {
  knitr::include_graphics(
    path = here::here(
      "Materials",
      data_source
    ),
    error = TRUE
  )
}

source(
  here::here("R/00_Config_file.R")
)

# Dynamic project name generation using {here}
project_name <- 
  basename(here::here())


# Create dynamic URLs
github_url <- 
  paste0(
    "https://github.com/OndrejMottl/",
     project_name
  )

slides_url <- 
  paste0(
    "https://ondrejmottl.github.io/", 
    project_name, "/"
  )
```

<br>

:::: {.columns}

::: {.column width="40%"}
:::

::: {.column width="20%"}

```{r}
#| label: personal sticker
include_local_figure("About/personal_sticker.png")
```

:::

::::

:::: {.r-hstack}
::: {data-id="box1" style="background: `r loqe_cols['teal']`; width: 135px; height: 15px; margin: 0px;"}
:::

::: {data-id="box2" style="background: `r loqe_cols['pistachio']`; width: 405px; height: 15px; margin: 0px;"}
:::

::: {data-id="box3" style="background: `r loqe_cols['seal_brown']`; width: 225px; height: 15px; margin: 0px;"}
:::

::: {data-id="box4" style="background: `r loqe_cols['fulvous']`; width: 135px; height: 15px; margin: 0px;"}
:::
::::

::: {.text-color-dark_green .text-font-heading .text-size-heading2 .text-bold .text-center}
Scientific profile of Ondřej Mottl
:::

<br>

:::: {.columns}

::: {.column width="20%"}

:::

::: {.column width="80%" .text-right .text-size-body}
[👤](https://ondrejmottl.github.io/) Ondřej Mottl

[MEMELAND Kick-Off 2025]{.text-highlight-pistachio}

:::

::::

<!--- This presentation --->

## [This presentation]{.text-color-white} {.title}

<br>

:::: {.columns}

::: {.column width="70%"}
::: {.r-fit-text}
[😸 Code on GitHub](`r github_url`)

[🖥️ Slides](`r slides_url`)

[![](https://img.shields.io/badge/License-MIT-yellow.svg)](https://opensource.org/licenses/MIT)
:::
:::

::: {.column width="30%"}
```{r}
#| label: qr_slides
qr_slides <-
  qrcode::qr_code(
    slides_url,
    ecl = "H"
  )

qrcode::generate_svg(
  qrcode = qr_slides,
  filename = here::here("Materials/QR/qr_slides.svg"),
  foreground = loqe_cols['white'],
  background = loqe_cols['dark_green']
)

include_local_figure("QR/qr_slides.svg")
```
:::

::::

::: {.text-center .text-size-heading3 .blockquote .text-color-black}
Scan to follow along!
:::

<!--- Slide 5: Hi, I am Ondra--->

## 🙋Hi! I am Ondřej Mottl {.text-center}

:::: {.columns}

::: {.column width="35%"}
```{r}
#| label: personal photo
include_local_figure("About/head_photo.png")
```

:::

::: {.column width="65%" .text-left}

<br>

🌿[Assistant Professor]{.text-highlight-dark_green} of [Vegetation Ecology]{.text-highlight-teal} at the [Department of Botany](https://natur.cuni.cz/en/biology/departments-and-work-places/department-of-botany)

🧑‍💻 Head of the [Laboratory of Quantitative Ecology](https://ondrejmottl.github.io/lab/about_the_lab.html)

♻️[Open Science Champion](https://openscience.cuni.cz/OSCIEN-161.html)

🎩Secretary of the [Junior Faculty](https://natur.cuni.cz/en/research/junior-faculty)

:::

::::

:::: {.columns}

::: {.column width="8%"}
:::

::: {.column width="60%" .text-left .text-smaller}

📧 Ondrej.mottl(at)gmail.com

🦋 [ondrejmottl.bsky.social](https://bsky.app/profile/ondrejmottl.bsky.social)

🐱 [Github](https://github.com/OndrejMottl)

🆔 [ORCID](https://orcid.org/0000-0002-9796-5081)

🌐 [Personal webpage](https://ondrejmottl.github.io/)
:::

::: {.column width="8%"}
:::


::: {.column width="22%"}

```{r}
#| label: linktree QR
qr_personal_web <-
  qrcode::qr_code(
    "https://bit.ly/ondrej_mottl",
    ecl = "H"
  )

qrcode::generate_svg(
  qrcode = qr_personal_web,
  filename = here::here("Materials/QR/qr_personal_web.svg"),
  foreground = loqe_cols['black'],
  background = loqe_cols['white']
)

include_local_figure("QR/qr_personal_web.svg")
```

:::

::: {.column width="5%"}
:::

::::

::: {.footer}
photo by Vladimír Šigut (CUNI)
:::


<!--- FOCUS --->

## Research Focus & Tools {.text-center}

<br>
<br>

:::: {.columns}

::: {.column width="30%" .fragment}
#### [Biodiversity patterns]{.text-highlight-dark_green} across scales

```{r}
#| label: biodiversity-patterns
include_local_figure("AI_generated/biodiversity_nobg.png")
```

:::

::: {.column width="3%"}
:::

::: {.column width="30%" .fragment}
#### [Big data]{.text-highlight-teal} & [methods advancement]{.text-highlight-teal}

```{r}
#| label: big-data
include_local_figure("AI_generated/big_data_nobg.png")
```

:::

::: {.column width="3%"}
:::

::: {.column width="30%" .fragment}
#### [Open Science]{.text-highlight-pistachio} & [reproducibility]{.text-highlight-pistachio}

```{r}
#| label: open-science
include_local_figure("AI_generated/open_science_nobg.png")
```

:::

::::

::: {.footer}
Images created by OpenAI GPT-5
:::

<!--- HOPE --->

# Humans on Planet Earth

<br>

:::: {.columns}
::: {.column width="75%"}

[HOPE]{.text-highlight-dark_green}: *Long-term impacts on biosphere dynamics* 

-  a European Research Council Advanced Grant project 
-  from 1 January 2018 to the end of 2023 (5.5-year)

:::

::: {.column width="25%" .text-right}

```{r}
#| label: hope-logo-1
include_local_figure("Logos/hope_logo.gif")
```

:::
::::

## HOPE team

```{r}
#| label: make all crop circles
#| echo: false
#| results: false

# John Birks
cropcircles::crop_circle(
  images = "https://upload.wikimedia.org/wikipedia/commons/2/27/Portrait_of_John_Birks_in_front_of_a_bookcase.jpg",
  to = here::here(
    "Materials",
    "Hope_team",
    "john_birks.jpg"
  )
  )

# John-Arvid Grytnes
cropcircles::crop_circle(
  images = "https://mountainsinmotion.w.uib.no/files/2022/04/Grytnes_cropped100-300x300.png",
  to = here::here(
    "Materials",
    "Hope_team",
    "john_arvid_grytnes.jpg"
  )
)

# Hilary Birks
cropcircles::crop_circle(
  images = "https://rse.org.uk/wp-content/uploads/2024/06/Hilary_Birks_0f03d69ebbc53e1b2d33545c0c5645f1-600x600.jpg",
  to = here::here(
    "Materials",
    "Hope_team",
    "hilary_birks.jpg"
  )
)

# Alistair Seddon
cropcircles::crop_circle(
  images = "https://bjerknessenteret-staging.ams3.digitaloceanspaces.com/people/picture-933-1473843470.gif",
  to = here::here(
    "Materials",
    "Hope_team",
    "alistair_seddon.jpg"
  )
)

# Suzette Flantua
cropcircles::crop_circle(
  images = "https://sflantua.github.io/profile_flantua.jpg",
  to = here::here(
    "Materials",
    "Hope_team",
    "suzette_flantua.jpg"
  )
)

# Kuber Bhata
cropcircles::crop_circle(
  images = "https://0.academia-photos.com/6921283/9757041/10957082/s200_kuber.bhatta.jpg",
  to = here::here(
    "Materials",
    "Hope_team",
    "kuber_bhata.jpg"
  )
)

# Ondřej Mottl
cropcircles::crop_circle(
  images = here::here(
    "Materials",
    "About",
    "ondrej_mottl.jpg"
  ),
  to = here::here(
    "Materials",
    "Hope_team",
    "ondrej_mottl.jpg"
  )
)

# Vivian Felde
cropcircles::crop_circle(
  images = "https://bjerknessenteret-staging.ams3.digitaloceanspaces.com/people/Felde_web.gif",
  to = here::here(
    "Materials",
    "Hope_team",
    "vivian_felde.jpg"
  )
)
```

:::: {.columns .text-center .text-smaller}

::: {.column width="25%"}

```{r}
#| label: john
include_local_figure("Hope_team/john_birks.jpg")
```

John Birks

```{r}
#| label: Suzette
include_local_figure("Hope_team/suzette_flantua.jpg")
```

Suzette Flantua

:::

::: {.column width="25%"}

```{r}
#| label: john arvid
include_local_figure("Hope_team/john_arvid_grytnes.jpg")
```

John-Arvid Grytnes


```{r}
#| label: kuber
include_local_figure("Hope_team/kuber_bhata.jpg")
```

Kuber Bhata

:::

::: {.column width="25%"}

```{r}
#| label: hilary
include_local_figure("Hope_team/hilary_birks.jpg")
```

Hilary Birks

```{r}
#| label: ondrej
include_local_figure("Hope_team/ondrej_mottl.jpg")
```

Ondřej Mottl

:::

::: {.column width="25%"}
```{r}
#| label: alistair
include_local_figure("Hope_team/alistair_seddon.jpg")
```

Alistair Seddon

```{r}
#| label: vivian
include_local_figure("Hope_team/vivian_felde.jpg")
```

Vivian Felde

:::

::::


<!--- FOSSILPOL --->

## FOSSILPOL

:::: {.columns}

::: {.column width = "50%"}
```{r}
#| label: fossilpol_logo_2
include_local_figure("Logos/FOSSILPOL_logo_600ppi.png")
```
:::

::: {.column width = "50%" .center-vertical .text-larger .text-center}
The workflow to process global [palaeoecological data]{.text-highlight-seal_brown} of fossil pollen for [vegetation]{.text-highlight-dark_green}-based [macroecological]{.text-highlight-teal} [synthesis]{.text-highlight-pistachio}
:::

::::

## The FOSSILPOL workflow

<br>

```{r}
#| label: workflow full
#| out.height: 100%
#| out.width: NULL
include_local_figure("Schematics/workflow_full.png")
```

::: {.footer}
Artwork by Milan Teunissen van Manen (@MilanTvM).
:::


## The FOSSILPOL workflow

:::: {.columns}

::: {.column width="75%" .fragment}

<br>

- Handles all the steps and flags places where the [user]{.text-highlight-teal} needs to specify their [choices]{.text-highlight-pistachio}
:::

::: {.column width="25%"}

```{r}
#| label: qr_fossilpol
qr_fossilpol <-
  qrcode::qr_code(
    "https://bit.ly/FOSSILPOL",
    ecl = "H"
  )

qrcode::generate_svg(
  qrcode = qr_fossilpol,
  filename = here::here("Materials/QR/qr_fossilpol.svg"),
  foreground = loqe_cols['dark_green'],
  background = loqe_cols['white']
)

include_local_figure("QR/qr_fossilpol.svg")
```

:::

::::

:::: {.columns .fragment}

::: {.column width="60%"}
- '*A guide to the processing and standardisation of global palaeoecological data for large-scale syntheses using fossil pollen*' (GEB 2023)
:::

::: {.column width="5%"}
:::

::: {.column width="35%"}

```{r}
#| label: gep_certificate
include_local_figure("Printscreens/GEB_certifiate.png")
```
:::

::::


<!--- RoC --->

## Rate of Vegetation Change (RoC)

<br>

:::: {.columns}

::: {.column width="25%"}

```{r}
#| label: rratepol_logo
include_local_figure("Logos/RRatepol_logo.png")
```

:::

::: {.column width="75%" .incremental}
- A robust R package to calculate the [Rate of Change (RoC)]{.text-highlight-seal_brown} in [palaeoecological]{.text-highlight-seal_brown} time series
- incorporates age uncertainties, standardisation of taxonomic richness, and multivariate dissimilarity coefficients

:::

::::

## Rate of Vegetation Change (RoC)

```{r}
#| label: science 1
include_local_figure("RoC/Figure1.jpeg")
```

_[Mottl, O.]{.text-highlight-black .bold} & Flantua, S.G.A._, Bhatta K.B., Felde, V.A., Giesecke, T., Goring, S., Grimm, E.C., Haberle, S., Hooghiemstra H., Ivory, S., Kuneš, P., Wolters, S., Seddon A.W.R, Williams J.W. (2021), [Global acceleration in rates of vegetation change over the past 18,000 years]{.text-highlight-teal .italic}. **Science** 372: 860-864.

## Rate of Vegetation Change (RoC)

```{r}
#| label: science 2
include_local_figure("RoC/Figure2.jpeg")
```



## Rate of Vegetation Change (RoC)

<br>

::: {.incremental}

- Rates of vegetation change during the Late Holocene are now as fast or faster than during the Last Deglaciation
- The start of the increased global rates of vegetation change is surprisingly early, about 3000-4000 years ago
- Ecosystem turnover rates have been rapidly accelerating over the last century, so our work suggests that modern acceleration is just the tip of the iceberg
- This is the first quantitative detection analysis of global biodiversity changes over the last 18,000 years, based on 1181 records from the Neotoma 
   + an amazing global community of scientists working together!

:::

<!--- BIODYNAMICS --->

# BIODYNAMICS

:::: {.columns}

::: {.column width="25%"}

```{r}
#| label: biodynamics-logo
include_local_figure("Logos/logo_BIODYNAMICS.png")
```

:::

::: {.column width="75%" .text-left}

<br>
<br>

Spatio-temporal dynamics of biotic interactions during the [Holocene]{.text-highlight-seal_brown}

:::

::::

[PI]{.text-highlight-fulvous} | 2023–2027 

Funded by the [Czech Science Foundation (GAČR)](https://gacr.cz/en/)

## VegVault

:::::: {.columns}

::::: {.column width="50%"}

```{r}
#| label: VegVault logo
include_local_figure("Logos/VegVault_logo.png")
```

:::::

::::: {.column width="50%" .incremental}

* A global SQLite database linking [interdisciplinary]{.text-highlight-black .text-color-white .text-bold} **plot-based** vegetation data! 
* Links to functional [traits]{.text-highlight-teal}, [climate]{.text-highlight-dark_green} variables, and soil properties 
* Built for exploring **biodiversity** dynamics across [time & space]{.text-highlight-seal_brown}
* Open Source & [Reproducible]{.text-highlight-pistachio}
* 🔗Learn more: [bit.ly/VegVault](https://bit.ly/VegVault)

:::: {.columns}

::: {.column width="30%"}
:::

::: {.column width="40%"}

```{r}
#| label: vegvault_qr_code
qr_vegvault <-
  qrcode::qr_code(
    "https://bit.ly/VegVault",
    ecl = "H"
  )

qrcode::generate_svg(
  qrcode = qr_vegvault,
  file = here::here("Materials/QR/qr_vegvault.svg"),
  foreground = loqe_cols['black'],
  background = loqe_cols['white']
)

include_local_figure("QR/qr_vegvault.svg")
```

:::

::: {.column width="30%"}
:::

::::

:::::

::::::


## VegVault

~110 GB | 32 tables & 87 variables | 480,000+ Datasets | 13M+ Samples | 110,000+ Taxa | 11M+ Trait values | 8 Abiotic variables

:::: {.columns .fragment}

#### Data sources 

::: {.column width="40%" .incremental}

- 🌿 [Paleoecological records]{.text-highlight-seal_brown} ([Neotoma]{.text-italic}, [FOSSILPOL]{.text-italic})
- 🌱 [Contemporary vegetation]{.text-highlight-dark_green} ([sPlotOpen]{.text-italic}, [BIEN]{.text-italic})
- 🌾 [Functional traits]{.text-highlight-teal} ([BIEN]{.text-italic}, [TRY]{.text-italic})
- 🌍 [Climate & soil data]{.text-highlight-pistachio} ([CHELSA]{.text-italic}, [WoSIS]{.text-italic})  
:::

::: {.column width="60%"}

```{r}
#| label: Scheme
include_local_figure("VegVault/assembly_viz_04.png")
```
:::

::::

# Functional Vegetation Paleoecology

:::: {.columns}

::: {.column width="25%"}

```{r}
#| label: fvp-logo
include_local_figure("Logos/logo_FVP.png")
```

:::

::: {.column width="75%" .text-left}

<br>

Using [past vegetation data]{.text-highlight-seal_brown} to anticipate climate impacts

Start up own independent group at Charles University

:::

::::

[PI]{.text-highlight-fulvous} | 2025–2029 

Funded by the [PRIMUS program of Charles University](https://cuni.cz/UKEN-1.html) 

## Laboratory of Quantitative Ecology {.text-center}

:::: {.columns}

::: {.column width="20%"}

<br>

```{r}
#| label: loqe-logo
include_local_figure("Logos/LoQE_logo.png")
```

<br>

```{r}
#| label: loqe-qr
qr_loqe <-
  qrcode::qr_code(
    "https://bit.ly/CUNI_QuantitativeEcology",
    ecl = "H"
  )

qrcode::generate_svg(
  qrcode = qr_loqe,
  filename = here::here("Materials/QR/qr_loqe.svg"),
  foreground = loqe_cols['dark_green'],
  background = loqe_cols['white']
)

include_local_figure("QR/qr_loqe.svg")
```


:::


::: {.column width="80%"}

```{r}
#| label: group_photo
include_local_figure("About/lab_photo_2025.jpg")
```


:::

::::

## Laboratory of Quantitative Ecology 

<br>

:::::: {.columns}

::::: {.column width="50%" .fragment}

:::: {.columns}

#### [PhD Students]{.text-highlight-dark_green}

::: {.column width="40%"}

```{r}
#| label: Bryan
include_local_figure("About/Bryan.jpg")
```

Bryan Novio

:::

::: {.column width="40%"}

```{r}
#| label: Naty
include_local_figure("About/Naty.png")
```

Natalie Hanusova

:::

::::

:::::

::::: {.column width="20%" .fragment}

#### [Technician]{.text-highlight-teal}

```{r}
#| label: Frieda
include_local_figure("About/Frieda.jpeg")
```

Friederika Wolke

:::::

::::: {.column width="20%" .fragment}

#### [Alumni]{.text-highlight-pistachio}

```{r}
#| label: Marketa
include_local_figure("About/Marketa.jpeg")
```

Marketa Tumova

:::::

::::::

::: {.footer}

[bit.ly/CUNI_QuantitativeEcology](https://bit.ly/CUNI_QuantitativeEcology)

:::

<!---PalaeOpen --->

# PalaeOpen

:::: {.columns}

::: {.column width="25%"}

```{r}
#| label: palaeopen-logo
include_local_figure("Logos/PalaeOpen_logo.png")
```

:::

::: {.column width="75%" .text-left}

<br>

Open [Palaeoecological Data]{.text-highlight-seal_brown} 

Organise workshops, training schools, and conferences for non-pollen paleoecologists

:::

::::

[Steering Committee Member]{.text-highlight-pistachio} | 2025-2029 

Now also ambassador of Neotoma database for whole Europe!

Funded by the [COST Action](https://www.cost.eu/actions/CA20129/)

# Open Science & Reproducibility

<br>

:::: {.columns}

::: {.column width="75%" .incremental}

- [Open Science Champion]{.text-highlight-pistachio} at Charles University
- Teaching [SPROuT course]{.text-highlight-teal} (Science Powered by Reproducibility, Openness and Teamwork) since 2024 at Charles University (2021-2023 at University of Bergen)
- Making [hands-on workshops]{.text-highlight-dark_green} for (non) academic institutions

:::

::: {.column width="25%"}

```{r}
#| label: open science champion
include_local_figure("Printscreens/Open_Science_Champion.png")
```

:::

::::

# Nerd and Tech stuff

:::: {.columns}

::: {.column width="15%"}

```{r}
#| label: logos
include_local_figure("Logos/RRatepol_logo.png")
include_local_figure("Logos/FOSSILPOL_logo_600ppi.png")
include_local_figure("Logos/vaultkeepr_logo.png")
```

:::

::: {.column width="5%"}
:::

::: {.column width="65%" .incremental}

<br>
<br>

- Large scale analysis of [big data]{.text-highlight-teal} (eg [reproducible pipelines]{.text-highlight-pistachio})
- Empowering scientific exploration through [user-friendly tools]{.text-highlight-dark_green}
   + Making [R packages]{.text-highlight-seal_brown}: {RRatepol}, {FOSSILPOL}, {vaultkeepr}, ...
- Making website and presentation using [QUARTO]{.text-highlight-fulvous}
- ... 

:::

::::

# MEMELAND - my role

<br>

::: {.incremental}

- [Data Management advisor]{.text-highlight-teal}
  + Data Management Plan (DMP), Project Communication Plan (PCP), etc.
- [Reproducible data analysis]{.text-highlight-pistachio} pipelines
   + Workflow documentation, version control, etc.
- Main [data analyst]{.text-highlight-dark_green} (biodiversity dynamics)
   + This requires the data to be [well managed and documented]{.text-highlight-seal_brown}!
- webmaster?

:::

<!--- Closing--->

## {.text-center}

<br>

:::: {.columns}

::: {.column width="20%"}

```{r}
#| label: loqe-logo-end
include_local_figure("Logos/LoQE_logo.png")
```

:::

::: {.column width="80%"}
:::

::::

:::: {.r-hstack}
::: {data-id="box1" style="background: `r loqe_cols['teal']`; width: 135px; height: 15px; margin: 0px;"}
:::

::: {data-id="box2" style="background: `r loqe_cols['pistachio']`; width: 405px; height: 15px; margin: 0px;"}
:::

::: {data-id="box3" style="background: `r loqe_cols['seal_brown']`; width: 225px; height: 15px; margin: 0px;"}
:::

::: {data-id="box4" style="background: `r loqe_cols['fulvous']`; width: 135px; height: 15px; margin: 0px;"}
:::
::::

<br>

::: {.text-color-dark_green .text-font-heading .text-bold .text-size-heading3}
Scientific profile of Ondřej Mottl
:::

<br>

:::: {.columns}

::: {.column width="20%"}

```{r}
#| label: loqe-qr-end
include_local_figure("QR/qr_loqe.svg")
```
:::

::: {.column width="60%" .text-left .text-size-body}
[👤](https://ondrejmottl.github.io/) Ondřej Mottl

[MEMELAND Kick-Off 2025]{.text-highlight-pistachio}
:::

::: {.column width="18%"}

```{r}
#| label: personal sticker 2 
include_local_figure("About/personal_sticker.png")
```

:::

::::
<!--- End of presentation --->